---
title: "Patterns"
author: "Ernesto Carrera"
date: "6/10/2017"
output:
  word_document: default
  html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Purpose:

The purpose of this document is to validate the relationship between the square
footage of the properties against their sales price.


We plot the variables in logarithmic scale to smooth the curve. However,
after doing the transformation there are several outliers in the data as the
plot below exhibits:

```{r pressure, echo=FALSE , results='hide',message=FALSE}

library(plyr)
library(gdata)
#setwd("C:/MSDS 6306-FALL2016/404/Live session 06")
#setwd("G://onlineSchool//SMU//MSDS6306//lectureNotes//week5")
setwd('/Users/ernestcarrera/Documents/Master Data Science/Doing Data Science/Session5/Assignment5/data')
getwd()



## You need a perl interpreter to do this on Windows.
## It's automatic in Mac
#bk <- read.xls("rollingsales_brooklyn.xls",pattern="BOROUGH")

# So, save the file as a csv and use read.csv instead
bk <- read.csv("rollingsales_brooklyn.csv",skip=4,header=TRUE)

## Check the data
#head(bk)
#summary(bk)
#str(bk) # Very handy function!

## clean/format the data with regular expressions
## More on these later. For now, know that the
## pattern "[^[:digit:]]" refers to members of the variable name that
## start with digits. We use the gsub command to replace them with a blank space.
# We create a new variable that is a "clean' version of sale.price.
# And sale.price.n is numeric, not a factor.
#ernesto's interp it is eliminating anything that is NOT A DIGIT.
#REGULAR EXPRESSION ARE SPECIFIED IN DOUBLE BRACKETS like this[[:digits:]]
#as.numeric coerces/creates/transform object to numeric and we record the result in SALE.PRICE.N
#count the number of missing observations
bk$SALE.PRICE.N <- as.numeric(gsub("[^[:digit:]]","", bk$SALE.PRICE))
count(is.na(bk$SALE.PRICE.N))

names(bk) <- tolower(names(bk)) # make all variable names lower case

## TODO: Get rid of leading digits bk$gross.square.feet as above bk$SALE.PRICE
bk$gross.sqft <- as.numeric(gsub("[^[:digit:]]","", bk$gross.square.feet))


# TODO: Get rid of leading digits of bk$land.sqft as above bk$SALE.PRICE
bk$land.sqft <- as.numeric(gsub("[^[:digit:]]","", bk$land.square.feet))
  
bk$year.built <- as.numeric(as.character(bk$year.built))

## do a bit of exploration to make sure there's not anything
## weird going on with sale prices
#attach(bk)
#hist(sale.price.n) 
#detach(bk)

```

```{r echo=FALSE , results='hide',message=FALSE}
bk.sale <- bk[bk$sale.price.n !=0,]
#We verify we do not have missing Values for the variables
summary(bk.sale$sale.price.n)
#create a LOGICAL vector with the indices that contain missing values
vect.bksalenotna <- complete.cases(bk$sale.price.n)  
str(vect.bksalenotna)
#using WHICH I GET RID OF THE MISSING VALUES
bk.sale1 <- bk.sale[which(vect.bksalenotna),]
summary(bk.sale$gross.sqft)
summary(bk.sale$sale.price.n)

#plot(bk.sale$gross.sqft,bk.sale$sale.price.n)
#plot(log10(bk.sale$gross.sqft),log10(bk.sale$sale.price.n))

```

```{r echo=FALSE , results='hide',message=FALSE}
## for now, let's look at 1-, 2-, and 3-family homes
bk.homes <- bk.sale1[which(grepl("FAMILY",bk.sale$building.class.category)),]
dim(bk.homes)


# TODO: complete plot() with log10 of bk.homes$gross.sqft,bk.homes$sale.price.n
#   as above "bk.sale"
plot(log10(bk.homes$gross.sqft),log10(bk.homes$sale.price.n))
summary(bk.homes[which(bk.homes$sale.price.n<100000),])


## remove outliers that seem like they weren't actual sales
bk.homes$outliers <- (log10(bk.homes$sale.price.n) <=5) + 0

# TODO: find out homes that meets bk.homes$outliers==0
bk.homes <- bk.homes[which(bk.homes$outliers==0),]
```


After removing the outliers we can see the exptected relationship
between gross square feet and Sale price. The higher the square feet
the higher the sale price
```{r echo=FALSE}
plot(log10(bk.homes$gross.sqft),log10(bk.homes$sale.price.n))
```

